A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-10-19 kobakian-cook.Rmd
Up to implementation

In many countries, residents are increasingly gathering to live in urban areas around major cities [@ACTUC]. The creation of approximately equal population areas results in dramatically different square meterage of the administrative geographic units, such as states or electorates. The units are filled with colour to represent the value of the statistic for each unit [@EI]. This can lead to unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for males for each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s around Brisbane stand out as more orange and red.

@@

@@

<!-- choropleth -->

The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

@@

```{r alternatives}
# Projected data 
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112) %>% filter(state_name_2011 == "Tasmania")
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "#D3D3D3", colour = "white", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + guides(fill=FALSE) +
  theme_map() + ggtitle("a")
#tas_ggcont
```

```{r contcart}
# Non - Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()
save(ncont, file = "data/ncont.rda")
}
```

```{r plotcontcart}
load("data/ncont.rda")
tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE) +
  theme_map() + ggtitle("b")
#tas_ggncont
```

```{r dorl}
# Non - Contiguous Dorling Cartograms
dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR)) + 
  scale_fill_identity() +
  guides(fill = FALSE) +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

@@

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:tas_displays}b) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS]. This shown in the northern SA2s of Tasmania with an average (yellow) estimated SIR of thyroid for males.

@@

The non-contiguous cartogram in figure \ref{fig:tas_displays}c) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart, shown in the inset map looks reasonable, but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in figure \ref{fig:tas_displays} a), or b).

@@

The purpose of this algorithm is to create a map display that highlights the spatial distributions for populations. There has been an increasing need for displays that recognise the large number of people that live in dense urban environments. The algorithm intends to maintain the spatial relationships of a group of geographic units in two ways: between each unit and its neighbours; and between each unit and the closest focal point. The algorithm allocates geographic units to a representative hexagon, in order of their proximity to the closest focal point.

The algorithm is named for the *Trigona carbonaria* bee species. Native to northern and eastern Australia this stingless species builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This hive design inspired the use of multiple focal points in the algorithm, individual spirals are constructed outward from various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely populated geographic regions surrounding the capital cities in Australia.
Each capital city forms it's own layer of connected hexagons. Figure \ref{fig:hexmap} allows the distributions in the capital cities to be viewed at a deeper scale. Sydney is now notable for the considerable amount of red areas, not just for the surrounding orange areas around the New South Wales capital visible in the choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also draw attention for the many orange areas. Interestingly there is no clear divide between rural and urban SA2 areas as Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis.

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for males, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap_plot
```

There are three key steps performed to create the tessellated hexagon tile map shown in Figure \ref{fig:hexmap}. First, derive the set of centroids from the polygons provided, then create the grid of hexagons locations. These two steps are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. To make tessellated plots, the point locations are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm, the shape file, and the id variable. The id variable should uniquely identify each geographic unit in the shape file.

The centroids can be derived from the shape file. The amount of centroids is used to determine a hexagon size. The derived centroids are a necessary input for an appropriate grid to be constructed, in which each grid point is located a hexagon size distance from the next closest point in all six directions. The grid will initially cover the entire map space, encompassing all the centroid locations and will extend in all directions to the extent of the buffer distance. This buffer distance can be helpful to account for densely populated coastal areas, allowing the use of the sea for hexagon locations. The hexagon size will likely need adjustments depending on the density of the population.

The centroids derived from the shape file are necessary inputs when creating a grid. The creation of the grid and filtering out the points unlikely to be used in the hexagon map results in the hexagon tile map grid. At this point, the centroid set and the grid become the necessary inputs to the allocation process.

A set of reference locations can be provided as focal points, such as capital cities of states or countries. Using focal points, the algorithm will create the order of allocation in order of the closest centroid locations to the focal points. Alternatively, a user can specify the variable that should be used to determine the order for the allocation. When allocating representative hexagons, the width parameter can be used to determine the flexibility of positioning using the relationship with the nearest reference location. A larger width parameter will increase the amount of available hexagon locations nearer to the centroid of the geographic location. A smaller width will maintain the orientation from the focal point to the centroid when selecting the hexagon location, however this could mean it is placed further from the centre of the cluster.

@@

Hexagon tile maps can be useful to understand a distribution across a collection of geographic areas. However, these maps are not easily drawn, especially as the number of areas increases. This algorithm was created to automate this process, and reduce the manual load involved in creating and implementing alternative displays. This allows map makers and data communicators to allocate their time to choosing the most effective display, rather than manually creating them.

@@

A user can tweak the parameters of the hexagon map using the additional arguments to the `create_hexmap` function, these options may affect the speed of the creation.

@@

## Hexagon tile map of Australia

This algorithm has been applied to the complete set of all Australian Statistical Areas at Level 2. This display highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for males in the communities of these three cities. There are several collections of red hexagons in Hobart that draw attention and they represent the communities with much higher rates of diagnosis than the Australian average. The communities south of Hobart show a gradient of colour as the values progress from higher than average, to much lower than average for the communities closer to the Sydney CBD. This pattern progresses into the communities with lower than average rates in Melbourne and Tasmania.

>>>>

Over time the density of living in major cities has increased as residents have gathered to live near urban areas [@ACTUC]. The division of these resident into approximately equal population areas results in dramatically different square meterage of administrative geographic units, such as states or electorates. In a choropleth map display the geographic units are filled with colour to represent the value of the statistic for each unit [@EI]. This can allow unequal emphasis on the statistical value of each of the geographic units. This allows choropleth map displays to misrepresent the spatial distributions of human related statistics due to area-size bias [@BCM]. Figure \ref{fig:choro} is a choropleth map that uses colour to display the estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for females, in each of the Statistical Areas at Level 2 (SA2) used by the @abs2011. The Australian choropleth map display draws attention to the expanse of dark and light blue areas across the rural communities in all states. The SA2s around Brisbane and in northern New South Wales stand out as more orange and red. However, this display neglects the vast amount of Australia residents living in the densely populated capital cities.

@@

<!-- choropleth -->
The familiar choropleth map display is often used to allow the user to orient themselves. However, the emphasis on the land mass diminishes the features of the distribution in densely populated communities due to the small size on the display [@ACTUC]. The unique shapes of boundaries can be helpful for orienting users but may not contribute to their understanding of the spatial disease distribution as many of the communities are not visible in a choropleth display [@TVSSS]. Each Statistical Area at Level 2 (SA2) [@abs2011] was designed to represent a community. These administrative areas are often used to aggregate census data and used to understand demographics within communities of the Australian population. This presents an opportunity for explicit map transformations to improve communication of distributions [@CBATCC]. In figure \ref{fig:tas_displays} the choropleth map can be seen underneath each alternative map display to allow for comparisons to be made.

@@

@@

```{r contcart}
# Projected data 
sa2thyroid_ERP <- st_transform(sa2thyroid_ERP, 3112) %>% filter(state_name_2011 == "Tasmania")
sa2thyroid_ERP <- sa2thyroid_ERP %>% 
  mutate(sva = sqrt(as.numeric(Population)/as.numeric(albers_sqkm)))

@@

tas_ggcont <- ggplot(cont) + 
  geom_sf(fill = "#D3D3D3", colour = "white", data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() + guides(fill=FALSE) +
  theme_map() + ggtitle("a")
#tas_ggcont
```

```{r ncontcart}
# Non - Contiguous Cartograms
if (!file.exists("data/ncont.rda")) {
ncont <- cartogram_ncont(sa2thyroid_ERP, k = 1.5,
                         weight = "Population") %>% 
                          st_as_sf()
save(ncont, file = "data/ncont.rda")}

load("data/ncont.rda")

tas_ggncont <- ggplot(ncont %>% filter(state_name_2011 == "Tasmania")) + 
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) +
  geom_sf(aes(fill = SIR), colour = NA) + 
  scale_fill_identity() + 
  guides(fill=FALSE) +
  theme_map() + ggtitle("b")
#tas_ggncont
```

```{r dorl}
# Non - Contiguous Dorling Cartograms
dorl <- sa2thyroid_ERP %>% 
  filter(!is.na(Population)) %>% 
  #mutate(pop = (Population/max(Population))*10) %>% # helpful for all of Australia
  cartogram_dorling(., weight = "Population", k = 0.2, m_weight = 1) %>% st_as_sf()
d <- st_bbox(dorl)
tas_ggdorl <- ggplot(dorl) +
  geom_sf(fill = "#D3D3D3", colour = "white",  data = sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), colour = "grey", size = 0.01) + 
  scale_fill_identity() +
  guides(fill = FALSE) +
  theme_map() + ggtitle("c")
#tas_ggdorl
```

@@

When communicating information that is relevant to the population, each member of the population can be given equal representation by transforming the map [@TVSSS]. The impact of a disease over connected communities can be seen by allowing the boundaries to maintain connection in the transformed display. The contiguous cartogram displayed in figure \ref{fig:tas_displays}a) draws attention to smaller geographic units when they are rescaled according to the population [@DMAHP]. These new shapes can now be coloured to represent a second variable. The algorithm uses the geographic shape of the areas and iterates toward sizing the areas to represent the population. This display can create twisted and unfamiliar shapes from the geographic units as the algorithms must satisfy the topology conditions, especially when there are communities located geographically far from their neighbours [@TVSSS].

@@

The non-contiguous cartogram in figure \ref{fig:tas_displays}b) also uses the population to rescale the geographic units. Unlike the contiguous cartogram, the SA2 areas maintain their geographic shape, but they may not retain the connection with their neighbours. The population of the SA2 areas is used to scale the geographic units in the non-contiguous cartogram [@NAC]. The amount of white space can be meaningful in non-contiguous cartograms [@ECGC], in this example the disparity between large rural areas and small urban areas means that the city of Hobart may look reasonable but these units are not visible in the context of the Tasmania state map. Depending on the difference between the population and geographic land size, the amount of white space can also prevent meaningful understanding of the distribution [@TVSSS].

@@

The Dorling cartogram presents each geographic unit as a circle, the size of the circle is scaled according to the population value of each area [@ACTUC]. Figure \ref{fig:tas_displays}c) shows the Tasmania SA2 areas as an individual circle located as close as possible to the geographic centroid location. This map draws attention to the collection of coastal cities in Tasmania that were not apparent in figure \ref{fig:tas_displays} a), or b). This display also highlights that the approximation of equal population is not easy to fulfil, as there is some disparity in the sizes of the circles.

@@

The purpose of this algorithm is to create an alternative map display that highlights the spatial distributions for populations. There has been an increasing need for displays that recognise the large number of people that live in dense urban environments. The algorithm intends to maintain the spatial relationships of a group of geographic units using the relationship between each unit and the closest focal point. The algorithm allocates geographic units to a representative hexagon, in order of their proximity to the closest focal point.

The algorithm is named for the *Trigona carbonaria* bee species. Native to northern and eastern Australia this stingless species builds flat layers of hexagonal brood cells, spiralling out from a central point [@PH]. This hive design inspired the use of multiple focal points in the algorithm, individual spirals are constructed by building out from various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely populated geographic regions surrounding the capital cities in Australia.
Each capital city forms it's own collection of connected hexagons. Figure \ref{fig:hexmap} allows the distributions in the capital cities to be viewed at a deeper scale. Sydney is now notable for the considerable amount of red hexagons, not just for the surrounding orange areas around the New South Wales capital visible in the choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also draw attention for the many orange hexagons. Interestingly there is no clear divide between rural and urban SA2 areas as many rural areas, and the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of thyroid cancer diagnosis for females.

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for females, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap_plot
```

There are three key steps performed to create the tessellated hexagon tile map shown in Figure \ref{fig:hexmap}. First, derive the set of centroids from the polygons provided, then create the grid of hexagons locations. These two processes are defined in the blue left column of the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be allocated to an available hexagon location. The steps for the allocation process are detailed in the right column of Figure \ref{fig:sugarbag_flow}. There are several filter steps to speed the process of selecting an appropriate hexagon to represent each geographic area. To make tessellated plots with the hexagon allocations, the point locations are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm in the `sugarbag` package; the shape file, and the id variable. The id variable should uniquely identify each geographic unit in the shape file.

The centroids are derived from the shape file. The amount of centroids within the geographic area is used to determine a hexagon size if one is not provided. The derived centroids are a necessary input for an appropriate grid to be constructed, each grid point is located a hexagon size distance from the next closest point in all six directions. The grid will initially cover the entire map space, encompassing all the centroid locations and will extend in all directions to the extent of the buffer distance. This buffer distance can be helpful to account for densely populated coastal areas, allowing the use of the sea for hexagon locations.

The centroid set derived from the shape file is necessary when creating a grid. The hexagon tile map grid requires the creation of the full grid followed by disregarding the points unlikely to be used in the hexagon map.

The centroid set and hexagon tile map grid are necessary for the allocation process. Additionally, a set of reference locations can be provided as focal points, such as the capital cities of states or countries. The algorithm will use the focal points to create the order of allocation, prioritising the closest centroid locations to the focal points. Alternatively, a user can specify the variable that should be used to determine the order for the allocation.

When allocating representative hexagons, the width parameter can be used to determine the flexibility of positioning. Using the relationship with the nearest reference location, a larger width parameter will increase the amount of available hexagon locations nearer to the centroid of the geographic location. A smaller width will maintain the orientation from the focal point to the centroid when selecting the hexagon location. However, this could mean it is placed further from the geographic centroid location.

@@

Hexagon tile maps can be useful to understand a distribution across a collection of geographic areas. However, these maps are not easy to create manually, especially as the number of areas increases. This algorithm was created to automate this process, and reduce the manual load involved in creating and implementing alternative displays. This allows map makers and data communicators to spend their time to choosing the most effective display.

@@

A user can tweak the parameters of the hexagon map using the additional arguments to the `create_hexmap` function, these options may affect the speed of the creation. 
The hexagon size may need adjustments depending on the density of the population, users can provide an appropriate hexagon size and re-run the algorithm. The buffer distance may need to be increased if the coastal cities need to extend beyond the geographic land mass.

@@

### Hexagon tile map of Australia

This algorithm has been applied to the complete set of all Australian Statistical Areas at Level 2. This display highlights the density of Australian capital cities, as it draws attention to the many communities in Sydney, Melbourne and Hobart. This display also highlights the disparity in the burden of thyroid cancer for females in the communities of these three cities. There are several collections of red hexagons in Hobart that draw attention and they represent the communities with much higher rates of diagnosis than the Australian average. The communities south of Hobart show a gradient of colour as the values progress from higher than average, to much lower than average for the communities closer to the Sydney CBD. This pattern progresses into the communities with lower than average rates in Melbourne and Tasmania.

### 2020-10-19 kobakian-cook.tex
Up to implementation

In many countries, residents are increasingly gathering to live in urban
areas around major cities \citep{ACTUC}. The creation of approximately
equal population areas results in dramatically different square meterage
of the administrative geographic units, such as states or electorates.
The units are filled with colour to represent the value of the statistic
for each unit \citep{EI}. This can lead to unequal emphasis on the
statistical value of each of the geographic units. This allows
choropleth map displays to misrepresent the spatial distributions of
human related statistics due to area-size bias \citep{BCM}. Figure
\ref{fig:choro} is a choropleth map that uses colour to display the
estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for
males for each of the Statistical Areas at Level 2 (SA2) used by the
\citet{abs2011}. The Australian choropleth map display draws attention
to the expanse of dark and light blue areas across the rural communities
in all states. The SA2s around Brisbane stand out as more orange and
red.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The impact of a disease over
connected communities can be seen by allowing the boundaries to maintain
connection in the transformed display. The contiguous cartogram
displayed in figure \ref{fig:tas_displays}b) draws attention to smaller
geographic units when they are rescaled according to the population
\citep{DMAHP}. These new shapes can now be coloured to represent a
second variable. This display can create twisted and unfamiliar shapes
from the geographic units as the algorithms must satisfy the topology
conditions, especially when there are communities located geographically
far from their neighbours \citep{TVSSS}. This shown in the northern SA2s
of Tasmania with an average (yellow) estimated SIR of thyroid for males.

The non-contiguous cartogram in figure \ref{fig:tas_displays}c) also
uses the population to rescale the geographic units. Unlike the
contiguous cartogram, the SA2 areas maintain their geographic shape, but
they may not retain the connection with their neighbours. The population
of the SA2 areas is used to scale the geographic units in the
non-contiguous cartogram \citep{NAC}. The amount of white space can be
meaningful in non-contiguous cartograms \citep{ECGC}, in this example
the disparity between large rural areas and small urban areas means that
the city of Hobart, shown in the inset map looks reasonable, but these
units are not visible in the context of the Tasmania state map.
Depending on the difference between the population and geographic land
size, the amount of white space can also prevent meaningful
understanding of the distribution \citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:tas_displays}c) shows the Tasmania
SA2 areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Tasmania that were not apparent in figure
\ref{fig:tas_displays} a), or b).

@@

The purpose of this algorithm is to create a map display that highlights
the spatial distributions for populations. There has been an increasing
need for displays that recognise the large number of people that live in
dense urban environments. The algorithm intends to maintain the spatial
relationships of a group of geographic units in two ways: between each
unit and its neighbours; and between each unit and the closest focal
point. The algorithm allocates geographic units to a representative
hexagon, in order of their proximity to the closest focal point.

The algorithm is named for the \emph{Trigona carbonaria} bee species.
Native to northern and eastern Australia this stingless species builds
flat layers of hexagonal brood cells, spiralling out from a central
point \citep{PH}. This hive design inspired the use of multiple focal
points in the algorithm, individual spirals are constructed outward from
various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely
populated geographic regions surrounding the capital cities in
Australia. Each capital city forms it's own layer of connected hexagons.
Figure \ref{fig:hexmap} allows the distributions in the capital cities
to be viewed at a deeper scale. Sydney is now notable for the
considerable amount of red areas, not just for the surrounding orange
areas around the New South Wales capital visible in the choropleth
display in Figure \ref{fig:choro}. Brisbane and Perth also draw
attention for the many orange areas. Interestingly there is no clear
divide between rural and urban SA2 areas as Melbourne, Darwin, Adelaide
and Hobart have low rates of thyroid cancer diagnosis.

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/hexmap-1} \caption[A hexagon tile map of the Statistical Areas of Australia at Level 2]{A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for males, they range from much lower than average (blue) to much higher than average (red).}\label{fig:hexmap}
\end{figure}
\end{Schunk}

There are three key steps performed to create the tessellated hexagon
tile map shown in Figure \ref{fig:hexmap}. First, derive the set of
centroids from the polygons provided, then create the grid of hexagons
locations. These two steps are defined in the blue left column of the
flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then be
allocated to an available hexagon location. The steps for the allocation
process are detailed in the right column of Figure
\ref{fig:sugarbag_flow}. To make tessellated plots, the point locations
are converted to hexagon shapes, with the hexagon size used in the grid.

@@

Only two inputs are necessary to begin using the algorithm, the shape
file, and the id variable. The id variable should uniquely identify each
geographic unit in the shape file.

The centroids can be derived from the shape file. The amount of
centroids is used to determine a hexagon size. The derived centroids are
a necessary input for an appropriate grid to be constructed, in which
each grid point is located a hexagon size distance from the next closest
point in all six directions. The grid will initially cover the entire
map space, encompassing all the centroid locations and will extend in
all directions to the extent of the buffer distance. This buffer
distance can be helpful to account for densely populated coastal areas,
allowing the use of the sea for hexagon locations. The hexagon size will
likely need adjustments depending on the density of the population.

The centroids derived from the shape file are necessary inputs when
creating a grid. The creation of the grid and filtering out the points
unlikely to be used in the hexagon map results in the hexagon tile map
grid. At this point, the centroid set and the grid become the necessary
inputs to the allocation process.

A set of reference locations can be provided as focal points, such as
capital cities of states or countries. Using focal points, the algorithm
will create the order of allocation in order of the closest centroid
locations to the focal points. Alternatively, a user can specify the
variable that should be used to determine the order for the allocation.
When allocating representative hexagons, the width parameter can be used
to determine the flexibility of positioning using the relationship with
the nearest reference location. A larger width parameter will increase
the amount of available hexagon locations nearer to the centroid of the
geographic location. A smaller width will maintain the orientation from
the focal point to the centroid when selecting the hexagon location,
however this could mean it is placed further from the centre of the
cluster.

@@

\hypertarget{hexagon-tile-map-of-australia}{%
\subsection{Hexagon tile map of
Australia}\label{hexagon-tile-map-of-australia}}

This algorithm has been applied to the complete set of all Australian
Statistical Areas at Level 2. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Hobart. This display also highlights the
disparity in the burden of thyroid cancer for males in the communities
of these three cities. There are several collections of red hexagons in
Hobart that draw attention and they represent the communities with much
higher rates of diagnosis than the Australian average. The communities
south of Hobart show a gradient of colour as the values progress from
higher than average, to much lower than average for the communities
closer to the Sydney CBD. This pattern progresses into the communities
with lower than average rates in Melbourne and Tasmania.

>>>>

Over time the density of living in major cities has increased as
residents have gathered to live near urban areas \citep{ACTUC}. The
division of these resident into approximately equal population areas
results in dramatically different square meterage of administrative
geographic units, such as states or electorates. In a choropleth map
display the geographic units are filled with colour to represent the
value of the statistic for each unit \citep{EI}. This can allow unequal
emphasis on the statistical value of each of the geographic units. This
allows choropleth map displays to misrepresent the spatial distributions
of human related statistics due to area-size bias \citep{BCM}. Figure
\ref{fig:choro} is a choropleth map that uses colour to display the
estimated Standardised Incidence Ratios (SIRs) of thyroid cancer for
females, in each of the Statistical Areas at Level 2 (SA2) used by the
\citet{abs2011}. The Australian choropleth map display draws attention
to the expanse of dark and light blue areas across the rural communities
in all states. The SA2s around Brisbane and in northern New South Wales
stand out as more orange and red. However, this display neglects the
vast amount of Australia residents living in the densely populated
capital cities.

@@

When communicating information that is relevant to the population, each
member of the population can be given equal representation by
transforming the map \citep{TVSSS}. The impact of a disease over
connected communities can be seen by allowing the boundaries to maintain
connection in the transformed display. The contiguous cartogram
displayed in figure \ref{fig:tas_displays}a) draws attention to smaller
geographic units when they are rescaled according to the population
\citep{DMAHP}. These new shapes can now be coloured to represent a
second variable. The algorithm uses the geographic shape of the areas
and iterates toward sizing the areas to represent the population. This
display can create twisted and unfamiliar shapes from the geographic
units as the algorithms must satisfy the topology conditions, especially
when there are communities located geographically far from their
neighbours \citep{TVSSS}.

The non-contiguous cartogram in figure \ref{fig:tas_displays}b) also
uses the population to rescale the geographic units. Unlike the
contiguous cartogram, the SA2 areas maintain their geographic shape, but
they may not retain the connection with their neighbours. The population
of the SA2 areas is used to scale the geographic units in the
non-contiguous cartogram \citep{NAC}. The amount of white space can be
meaningful in non-contiguous cartograms \citep{ECGC}, in this example
the disparity between large rural areas and small urban areas means that
the city of Hobart may look reasonable but these units are not visible
in the context of the Tasmania state map. Depending on the difference
between the population and geographic land size, the amount of white
space can also prevent meaningful understanding of the distribution
\citep{TVSSS}.

The Dorling cartogram presents each geographic unit as a circle, the
size of the circle is scaled according to the population value of each
area \citep{ACTUC}. Figure \ref{fig:tas_displays}c) shows the Tasmania
SA2 areas as an individual circle located as close as possible to the
geographic centroid location. This map draws attention to the collection
of coastal cities in Tasmania that were not apparent in figure
\ref{fig:tas_displays} a), or b). This display also highlights that the
approximation of equal population is not easy to fulfil, as there is
some disparity in the sizes of the circles.

@@

The purpose of this algorithm is to create an alternative map display
that highlights the spatial distributions for populations. There has
been an increasing need for displays that recognise the large number of
people that live in dense urban environments. The algorithm intends to
maintain the spatial relationships of a group of geographic units using
the relationship between each unit and the closest focal point. The
algorithm allocates geographic units to a representative hexagon, in
order of their proximity to the closest focal point.

The algorithm is named for the \emph{Trigona carbonaria} bee species.
Native to northern and eastern Australia this stingless species builds
flat layers of hexagonal brood cells, spiralling out from a central
point \citep{PH}. This hive design inspired the use of multiple focal
points in the algorithm, individual spirals are constructed by building
out from various points on the geographic map base.

The result of the algorithm is a display that emphasises the densely
populated geographic regions surrounding the capital cities in
Australia. Each capital city forms it's own collection of connected
hexagons. Figure \ref{fig:hexmap} allows the distributions in the
capital cities to be viewed at a deeper scale. Sydney is now notable for
the considerable amount of red hexagons, not just for the surrounding
orange areas around the New South Wales capital visible in the
choropleth display in Figure \ref{fig:choro}. Brisbane and Perth also
draw attention for the many orange hexagons. Interestingly there is no
clear divide between rural and urban SA2 areas as many rural areas, and
the cities of Melbourne, Darwin, Adelaide and Hobart have low rates of
thyroid cancer diagnosis for females.

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/hexmap-1} \caption[A hexagon tile map of the Statistical Areas of Australia at Level 2]{A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for females, they range from much lower than average (blue) to much higher than average (red).}\label{fig:hexmap}
\end{figure}
\end{Schunk}

There are three key steps performed to create the tessellated hexagon
tile map shown in Figure \ref{fig:hexmap}. First, derive the set of
centroids from the polygons provided, then create the grid of hexagons
locations. These two processes are defined in the blue left column of
the flow chart in Figure \ref{fig:sugarbag_flow}. Each centroid can then
be allocated to an available hexagon location. The steps for the
allocation process are detailed in the right column of Figure
\ref{fig:sugarbag_flow}. There are several filter steps to speed the
process of selecting an appropriate hexagon to represent each geographic
area. To make tessellated plots with the hexagon allocations, the point
locations are converted to hexagon shapes, with the hexagon size used in
the grid.

@@

Only two inputs are necessary to begin using the algorithm in the
\texttt{sugarbag} package; the shape file, and the id variable. The id
variable should uniquely identify each geographic unit in the shape
file.

The centroids are derived from the shape file. The amount of centroids
within the geographic area is used to determine a hexagon size if one is
not provided. The derived centroids are a necessary input for an
appropriate grid to be constructed, each grid point is located a hexagon
size distance from the next closest point in all six directions. The
grid will initially cover the entire map space, encompassing all the
centroid locations and will extend in all directions to the extent of
the buffer distance. This buffer distance can be helpful to account for
densely populated coastal areas, allowing the use of the sea for hexagon
locations.

The centroid set derived from the shape file is necessary when creating
a grid. The hexagon tile map grid requires the creation of the full grid
followed by disregarding the points unlikely to be used in the hexagon
map.

The centroid set and hexagon tile map grid are necessary for the
allocation process. Additionally, a set of reference locations can be
provided as focal points, such as the capital cities of states or
countries. The algorithm will use the focal points to create the order
of allocation, prioritising the closest centroid locations to the focal
points. Alternatively, a user can specify the variable that should be
used to determine the order for the allocation.

When allocating representative hexagons, the width parameter can be used
to determine the flexibility of positioning. Using the relationship with
the nearest reference location, a larger width parameter will increase
the amount of available hexagon locations nearer to the centroid of the
geographic location. A smaller width will maintain the orientation from
the focal point to the centroid when selecting the hexagon location.
However, this could mean it is placed further from the geographic
centroid location.

@@

\hypertarget{hexagon-tile-map-of-australia}{%
\subsubsection{Hexagon tile map of
Australia}\label{hexagon-tile-map-of-australia}}

This algorithm has been applied to the complete set of all Australian
Statistical Areas at Level 2. This display highlights the density of
Australian capital cities, as it draws attention to the many communities
in Sydney, Melbourne and Hobart. This display also highlights the
disparity in the burden of thyroid cancer for females in the communities
of these three cities. There are several collections of red hexagons in
Hobart that draw attention and they represent the communities with much
higher rates of diagnosis than the Australian average. The communities
south of Hobart show a gradient of colour as the values progress from
higher than average, to much lower than average for the communities
closer to the Sydney CBD. This pattern progresses into the communities
with lower than average rates in Melbourne and Tasmania.
